---
title: "Supplementary Code"
subtitle: "Bycatch mitigation could prevent strong changes in the ecological strategies of seabird communities across the globe"
author: "Cerren Richards"
date: '15-04-2021'
output: word_document
---

```{r setup, include=FALSE}

library(rmarkdown)
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '/Users/cerren/Desktop/Trait shifts - GECCO/Revisions/Seabird-Trait-Shifts')

```

#___________________________________________________________________
## Code Contents

1. Preparing BirdLife International species distribution polygons  
    * Import and subset the seabird polygons 
    * Create a species presence-absence matrix (PAM)
    * Remove the land from the PAM
    
2. Preparing the Global Fishing Watch data 
    * Extract latitude and longitude for the fishing gears largely responsible for seabird bycatch
    * Remove cells that are overlapping between fisheries and threatened species

3. Importing species' threat data from the IUCN
    * Subset species at risk to bycatch
    
4. Remove cells that are overlapping between fisheries and threatened species
    * Create global map of fishing distribution
    
5. Calculate the community weighted mean (CWM) metric
    * Observed CWM
    * CWM after removing bycatch-threatened species
    * Calculate the shift in CONTINUOUS traits
    * Plotting the shift in CONTINUOUS traits with GAM and qGAM
    * Calculate the shift in foraging guild (CATEGORICAL)
    * Plot the shift in foraging guild (CATEGORICAL) with GAMM
    * Calculate the shift in diet guild (CATEGORICAL)
    * Plot the shift in diet guild (CATEGORICAL) with GAMM


#___________________________________________________________________


## **1. Preparing BirdLife International species distribution polygons**

We downloaded the Bird of the World polygons from BirdLife International. These spatial polygons represent the coarse distributions that species likely occupy, and are presently the best available data for the ranges of all seabirds.

Bird distribution data are available upon request from http://datazone.birdlife.org/species/requestdis/.

### *Import and subset the seabird polygons* 

Here we subset the seabird polygons based on the list of seabirds from Richards et al. (2021) - Biological traits of seabirds predict extinction risk and vulnerability to anthropogenic threats - https://doi.org/10.1111/geb.13279

Imputed trait data are available for download from Dryad https://doi.org/10.5061/dryad.x69p8czhd. 


```{r message=FALSE, error=FALSE, warning=FALSE}
library(rgdal); library(dplyr)

# Read in the Bird of the World (BOTW) data [this takes a long time]
BOTW <- readOGR(dsn = "/Users/cerridwenr/Desktop/Masters Thesis /Chapter two/BOTW.gdb", layer = "All_Species")

# Read in the trait and seabird data
traits <- read.csv("Imputed_Trait_Data.csv")


# read in seaduck data
duck <- read.csv("seaducks.csv")
duck$body_mass_median <- log10(duck$body_mass_median)
duck$GL <- log10(duck$GL)

# Join the seabird and sea duck dataframes
traits <- rbind(traits, duck)

# BirdLife International does not have the distribution layers for Thayer's gull, so remove from analysis
traits <- traits %>% filter(binomial != "Larus thayeri")


# Rearrange the names based on alphabetical order
traits <- traits %>% arrange(binomial)


# Subset the BOTW dataset to only contain seabirds
sb<- subset(BOTW, SCINAME %in% traits$Binomial)

#save the SHAPE file as an RDS
saveRDS(sb, "BOTW_seabirds_361.rds")

```


### *Create a species presence-absence matrix (PAM)*

Here we further subset the seabird polygons to only retain the extant, native, resident, breeding season and non-breeding season polygons. 

Second, we create a 1° resolution global presence-absence matrix based on the seabird distribution polygons.

For more information on the codes used within the distribution polygons visit: http://datazone.birdlife.org/species/spcdistPOS


```{r message=FALSE, error=FALSE, warning=FALSE}
library(letsR)

# .rds file that contains the subsetted seabird polygons from the large BOTW data file 
sb <- readRDS("BOTW_seabirds_341.rds") # note this file has different binomials from the NEW_TRAITS file

# Only keep the breeding season, non-breeding season and resident polygons
# 1 = resident; 2 = breeding; 3 = non-breeding
seabirds <- sb[sb$SEASONAL %in% c(1, 2, 3), ]

# Only keep species that are extant
# 1 = Extant 
seabirds <- seabirds[seabirds$PRESENCE %in% c(1), ]

# Only keep species that are native
# 1 = Native
seabirds <- seabirds[seabirds$ORIGIN %in% c(1), ]

# Create PAM. This take ~30-40 minutes for 341 species
# Returns latitude and longitude with PAM for each species 
# Lat long values returned are the 1 degree cell's centroid
PAM <- lets.presab(seabirds, resol = 1, xmn = -180, xmx = 180, ymn = -90, ymx = 90, count = TRUE)

# Save PAM as a .rds file
saveRDS(PAM, "Polygon_PAM.rds")

```


### *Remove the land from the PAM*

We remove all land from the PAM because in analyses below we will be calculating ocean trait shifts, and the land obscures the pattern. 

```{r}
library(maptools); library(dplyr) 

data(wrld_simpl) # Extract polygon of the world

# This crops all cells from the ocean 
PAM_crop <- lets.pamcrop(PAM, wrld_simpl, remove.sp = TRUE)

# Plot to see
plot(PAM_crop, xlab = "Longitude", ylab = "Latitude",
     col = colorRampPalette(c("darkgreen", "yellow", "blue")))

# Save the PAM for the land 
PAM_crop_df <- as.data.frame(PAM_crop[["Presence_and_Absence_Matrix"]])

# Save the complete PAM (land & ocean)
PAM_df <- as.data.frame(PAM[["Presence_and_Absence_Matrix"]])

# Remove the land data from the complete PAM_df
PAM_ocean <-  anti_join(PAM_df, PAM_crop_df, by=c("Longitude(x)", "Latitude(y)"))

# Check whether it has successfully removed the land
plot(PAM_ocean$`Longitude(x)`, PAM_ocean$`Latitude(y)`)

saveRDS(PAM_ocean, "PAM_ocean.rds")

```



## **2. Preparing the Global Fishing Watch data** 

Here we use the 2015 - 2018 Global Fishing Watch data. These data were provided as the monthly sum of fishing hours per gear type per 1x1 degree cell. The data points are located in the southwestern corner of a respective cell. To centre the data within the cell, all points will need to be shifted by 0.5 degree North and East.

Fishing effort is available for the time period 2012 to 2016 from https://globalfishingwatch.org/. To access to provisional 2017 and 2018 data, contact research@globalfishingwatch.org.

### *Extract latitude and longitude for the fishing gears largely responsible for seabird bycatch*

Longlines, trawlers, purse seines and gill nets are the fishing gears responsible for the majority of seabird bycatch. 

In this section, we subset the Global Fishing Watch data to these four gears and remove any duplicate lat longs to produce a single map of the distribution for these gear types.

The fishing effort .csv files are very large, so we read each year in individually and reduce the data to prevent R from crashing.

```{r}
### 2015 ###

# read in the 2015 data 
fish15 <- read.csv("MUN_fishing_effort_2015.csv")

# Make the lat and long in the correct format
fish15$lat_bin <- (fish15$lat_bin)/100
fish15$lon_bin <- (fish15$lon_bin)/100

# Shift the lat long to the center of cells
fish15$lat_bin <- (fish15$lat_bin)+0.5
fish15$lon_bin <- (fish15$lon_bin)+0.5

# Subset the data to only the four gears largely responsible for seabird bycatch
fish15 <- filter(fish15, geartype %in% c("drifting_longlines", "set_longlines",
                                                "trawlers",
                                                "other_purse_seines", "purse_seines", "tuna_purse_seines",
                                                "set_gillnets"))

# only keep distinct lat longs - remove duplicates
fish15 <- distinct(fish15, lat_bin, lon_bin)

saveRDS(fish15, "fish15.rds")

# Clear the R environment 
rm(list = ls())




### 2016 ###


# read in the 2016 data 
fish16 <- read.csv("MUN_fishing_effort_2016.csv")

# Make the lat and long in the correct format
fish16$lat_bin <- (fish16$lat_bin)/100
fish16$lon_bin <- (fish16$lon_bin)/100

# Shift the lat long to the center of cells
fish16$lat_bin <- (fish16$lat_bin)+0.5
fish16$lon_bin <- (fish16$lon_bin)+0.5

# Subset the data to only the four gears largely responsible for seabird bycatch
fish16 <- filter(fish16, geartype %in% c("drifting_longlines", "set_longlines",
                                         "trawlers",
                                         "other_purse_seines", "purse_seines", "tuna_purse_seines",
                                         "set_gillnets"))

# only keep distinct lat longs - remove duplicates
fish16 <- distinct(fish16, lat_bin, lon_bin)

saveRDS(fish16, "fish16.rds")

# Clear the R environment 
rm(list = ls())





### 2017 ###

# read in the 2017 data 
fish17 <- read.csv("MUN_fishing_effort_2017.csv")

# Make the lat and long in the correct format
fish17$lat_bin <- (fish17$lat_bin)/100
fish17$lon_bin <- (fish17$lon_bin)/100

# Shift the lat long to the center of cells
fish17$lat_bin <- (fish17$lat_bin)+0.5
fish17$lon_bin <- (fish17$lon_bin)+0.5


# Subset the data to only the four gears largely responsible for seabird bycatch
fish17 <- filter(fish17, geartype %in% c("drifting_longlines", "set_longlines",
                                         "trawlers",
                                         "other_purse_seines", "purse_seines", "tuna_purse_seines",
                                         "set_gillnets"))

# only keep distinct lat longs - remove duplicates
fish17 <- distinct(fish17, lat_bin, lon_bin)

saveRDS(fish17, "fish17.rds")

# Clear the R environment 
rm(list = ls())





### 2018 ###


# read in the 2018 data 
fish18 <- read.csv("MUN_fishing_effort_2018.csv")

# Make the lat and long in the correct format
fish18$lat_bin <- (fish18$lat_bin)/100
fish18$lon_bin <- (fish18$lon_bin)/100

# Shift the lat long to the center of cells
fish18$lat_bin <- (fish18$lat_bin)+0.5
fish18$lon_bin <- (fish18$lon_bin)+0.5

# Subset the data to only the four gears largely responsible for seabird bycatch
fish18 <- filter(fish18, geartype %in% c("drifting_longlines", "set_longlines",
                                         "trawlers",
                                         "other_purse_seines", "purse_seines", "tuna_purse_seines",
                                         "set_gillnets"))

# only keep distinct lat longs - remove duplicates
fish18 <- distinct(fish18, lat_bin, lon_bin)

saveRDS(fish18, "fish18.rds")



#### Join all of the fishing years together and remove duplicate lat longs if present
fish18 <- readRDS("fish18.rds")
fish15 <- readRDS("fish15.rds")


tot_fish <- rbind(fish15, fish18)
remove(fish15, fish18)
tot_fish <- distinct(tot_fish, lat_bin, lon_bin)


fish16<- readRDS("fish16.rds")
tot_fish <- rbind(tot_fish, fish16)
remove(fish16)
tot_fish <-  distinct(tot_fish, lat_bin, lon_bin)


fish17 <- readRDS("fish17.rds")
tot_fish <- rbind(tot_fish, fish17)
remove(fish17)
tot_fish <-  distinct(tot_fish, lat_bin, lon_bin)


# Save the final .rds that contains the lat and longs for the four gears largely responsible for seabird bycatch 
saveRDS(tot_fish, "total_fish_distribution.rds")


```




## **3. Importing species' threat data from the IUCN**

To identify which species are threatened from bycatch, we first must extract all the seabird threats from the IUCN Threat Classification Scheme: https://www.iucnredlist.org/resources/threat-classification-scheme. 

IUCN Threat Data were downloaded on July 14th, 2020. Because threats are updated through time, they may change in the future.

Note - an API key is needed to download the IUCN data. APIs can be requested from: https://apiv3.iucnredlist.org/api/v3/token

```{r message=FALSE, error=FALSE, warning=FALSE, eval = FALSE}
library(rredlist); library(rlist) # Load package

spp <- as.character(traits$binomial) # vector of species names from the traits dataframe

iucn_key <- " " # Enter your unique IUCN code here

# Extract the threats for each species from th IUCN database. 
# NOTE - this code take ~15 minutes to run.
iucn <- lapply(spp, function(x) {
  y <- rl_threats(name = x, key = iucn_key)
  Sys.sleep(2)
  # 2 second delay makes API work better - recommended by IUCN
  return(y)
})

### Add an extra column to the dataframes in the list with the binomial names:
for (i in 1:length(iucn)) { 
  iucn[[i]][["result"]]$binomial <- iucn[[i]][["name"]] 
}

iucn <- Reduce(rbind, iucn) ## Reduce the lists down

iucn[1:361] <- NULL # remove all of the extra species names in the list

# Remove all lists with species with no threats
iucn <- iucn[sapply(iucn, length)>1] 

# binds all list elements by row to make a dataframe
iucn <- list.rbind(iucn) 

# Delete the columns that will not be used in further analyses
iucn <- iucn[- c(3:7)] 
```


### *Subset species at risk to bycatch*

Here we subset all the seabird species threatend from bycatch.

IUCN Codes = https://www.iucnredlist.org/resources/threat-classification-scheme

```{r, eval = FALSE}
# Subset based on the IUCN codes indicating unintentional effects of fishing
bycatch <- iucn[iucn$code %in% c("5.4.3", "5.4.4"), ]

#remove duplicates
bycatch <- bycatch[!duplicated(bycatch$binomial), ] 
bycatch <- as.data.frame(bycatch$binomial); colnames(bycatch) <- "Species"
bycatch$bycatch <- 1

write.csv(bycatch, "BycatchThreatendSeabirds.csv")
```


## **4. Remove cells that are overlapping between fisheries and threatened species**

Here we ensure that the seabird and vessel raster align by reprojecting them.

We then remove any areas where seabirds and vessels overlap for later analyses.

```{r}
library(raster)

vessels <- readRDS("total_fish_distribution.rds")
birds <- readRDS("PAM_ocean.rds")
threatened <- read.csv("BycatchThreatendSeabirds.csv")

##### VESSEL DATA #####

## Add a "z" column to the vessel data to run the `rasterFromXYZ()` code below
vessels$z <- 1

## Reorder the vessel columns to be in the order xyz
vessels <- vessels %>% dplyr::select(lon_bin, lat_bin, z)


# Note - the `vessels` file contains over 53 million data points
# The following code takes a LONG time to run

## Project fishing vessel and seabird distributions into rasters
vesselRaster <- rasterFromXYZ(vessels)

#lat and lon are in 0.1 units, so raster resolution is 0.01
# we will reproject the raster to 1x1 degree
vesselRaster <- aggregate(vesselRaster,fact=100,fun=max)
vesselRaster 
plot(vesselRaster)
projection(vesselRaster) <- crs("+proj=longlat +datum=WGS84")
res(vesselRaster)


##### BIRD DATA #####

## Add a column to indicate whether any threatened species is present
birds$all <- apply(birds[,3:363],1,max) 

#project bird data
birdRaster <- rasterFromXYZ(birds[,c("Longitude(x)","Latitude(y)","all")])
plot(birdRaster) 
projection(birdRaster) <- crs("+proj=longlat +datum=WGS84")
birdRaster 

## Reproject the vessel layer into the same as the bird layer
vesselRaster <- projectRaster(vesselRaster,birdRaster) 
plot(vesselRaster)

## make both a data frame and merge
vesselDF <- as.data.frame(vesselRaster,xy=T)
birdDF <- as.data.frame(birdRaster,xy=T)
allDF <- left_join(birdDF,vesselDF)


###### Joining the vessel data to the bird data ####

## Filter only the species threatened from bycatch
bycatch_birds <- birds %>% dplyr::select(`Longitude(x)`, `Latitude(y)`, threatened$binomial)

# Rename the columns of `vesselDF` to match the `bycatch_birds` columns
colnames(vesselDF) <- c("Longitude(x)", "Latitude(y)", "vessels")

# Join the two datasets
bycatch_birds <- left_join(bycatch_birds, vesselDF)

# Replace NAs in the `vessels` column with zeros
bycatch_birds$vessels[is.na(bycatch_birds$vessels)] <- 0

## Remove locations where birds and vessels overlap
## This is achieved through subtracting the vessel column (149) from all seabird columns (3:148)
for(i in colnames(bycatch_birds[,3:148])) {
  bycatch_birds[i] <- bycatch_birds[i] - bycatch_birds[149]
}

## Ensure all the values are rounded to 0 decimal places
birdround <- round(bycatch_birds[3:148], digits = 0)

## Any zero bird values have now been changed to `-1`. 
## We will now replace all `-1` values with zero
birdround[birdround == -1] <- 0

## Add the latitude and longitude back into the dataframe because it was removed during the rounding
## We will also add back the species that are not threatened from bycatch

## Drop all of the threatened bird column
birds_final <- dplyr::select(birds, -threatened$binomial)

## bind the threatened birds with the non threatened bird data
birds_final <- bind_cols(birds_final, birdround)

## I WILL NEED TO REORDER THESE COLUMNS TO ALPHABETIC
## https://stackoverflow.com/questions/29873293/dply-order-columns-alphabetically-in-r

## Save the data
saveRDS(birds_final, "PAM_No_Threatened_Overlap.rds")


## Comparing the seabird rasters after the overlapping areas are removed
full <- birds %>% dplyr::select(`Longitude(x)`, `Latitude(y)`, threatened$binomial)
removed <- birds_final %>% dplyr::select(`Longitude(x)`, `Latitude(y)`, threatened$binomial)

full$all <- apply(full[,3:148],1,max) 
removed$all <- apply(removed[,3:148],1,max)

# ORIGIONAL BIRD DATA
fullraster <- rasterFromXYZ(full[,c("Longitude(x)","Latitude(y)","all")])
plot(fullraster) 

# OVERLAP REMOVED
removedraster <- rasterFromXYZ(removed[,c("Longitude(x)","Latitude(y)","all")])
plot(removedraster) 


```

### *Create global map of fishing distribution*

Map the distributions of all four gears combined

```{r}
library(ggplot2);library(viridis); library(ggpubr); library(tidyverse); 
library(sf);library("rnaturalearth"); library("rnaturalearthdata")

# Extract only the cells that have fishing activity 
vesselDF2 <- vesselDF %>% filter(vessels == 1)

# Extract world polygon
world <- ne_countries(scale = "medium", returnclass = "sf")

# Plot
ggplot() +
  theme_bw(base_size = 18)+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  geom_tile(aes(x = x, y = y), fill = "darkturquoise", colour = "darkturquoise",data = vesselDF2) + 
  scale_y_continuous(expand = c(0,0), limits = c(-84, 85))+
  scale_x_continuous(expand = c(0,0), limits = c(-180, 180))+
  theme(legend.position = "none")+
  theme(axis.title.x=element_blank(), axis.title.y=element_blank(),)+
  geom_sf(data = world, fill = "grey60", colour = "grey20", size = 0.25)

ggsave("fishing_vessels.png")

```



## **5. Calculate the Community Weighted Mean (CWM) metric**
### *Observed CWM*

Run this with all the species present (n = 341). 

```{r message=FALSE, error=FALSE, warning=FALSE}
library(FD)

PAM <- readRDS("PAM_ocean.rds")

# Duplicate the dataframe because we will use the columns again
PAM2 <- PAM

#rownames(PAM) <- PAM$Locations
PAM$`Longitude(x)` <- NULL
PAM$`Latitude(y)` <- NULL
PAM <- as.matrix(PAM)

# Read in traits df and reformat it
list.files()
traits <- read.csv("Imputed Trait Data.csv")
traits$Order <- NULL
traits$Family <- NULL
traits$family_english <- NULL
traits$English <- NULL
traits$IUCN <- NULL
traits$Threat <- NULL

#set up dataframe in correct layout
rownames(traits) <- colnames(PAM)
traits$binomial <- NULL

# Keep only the traits to be used in the study
traits$hab_breadth <- NULL
traits$migrate <- NULL
traits$pelagic_specialist <- NULL

# Calculate the community weighted mean
CWM_all <- functcomp(traits, PAM, CWM.type = "dom", bin.num = NULL) 

# Rejoin the lat and long columns
CWM_all$Lat <- PAM2$`Latitude(y)`
CWM_all$Long <- PAM2$`Longitude(x)`

# Save
saveRDS(CWM_all, "CWM_all.rds")

# Read in if closed R
CWM_all <- readRDS("CWM_all.rds")

```


### *CWM after removing bycatch-threatened species* 

Recalculate the CWM with the PAM that had overlap with fishing removed

```{r}

PAM_overlap <- readRDS("PAM_No_Threatened_Overlap.rds")
PAM_overlap$all <- NULL

# Reorder the columns of `PAM_overlap` to match the PAM2 column orders
# this is important for caluclating the CWM because the orders of the species
# must match the traits order
PAM_overlap <- PAM_overlap[names(PAM2)]

# Remove the lat and long columns because we need to sum the rows below
PAM_overlap$`Longitude(x)` <- NULL
PAM_overlap$`Latitude(y)` <- NULL

# Some cells now have no species, so we must remove these before calculating the CWM
# also need to make sure to rematch up the latlongs
PAM_overlap$freq <- rowSums(PAM_overlap) 
PAM_overlap$Long <- PAM2$`Longitude(x)`
PAM_overlap$Lat <- PAM2$`Latitude(y)`
PAM_overlap <- PAM_overlap %>% filter(freq != 0)
PAM_overlap_nona <- PAM_overlap # for lat long reference

PAM_overlap$freq <- NULL
PAM_overlap$Lat <- NULL
PAM_overlap$Long <- NULL

# Turn `PAM_overlap` into a matrix
PAM_overlap <- as.matrix(PAM_overlap)

# Calculate the community weighted mean
CWM_overlap <- functcomp(traits, PAM_overlap, CWM.type = "dom", bin.num = NULL)

#add lat longs back
CWM_overlap$Long <- PAM_overlap_nona$Long
CWM_overlap$Lat <-  PAM_overlap_nona$Lat

# Save
saveRDS(CWM_overlap, "CWM_overlap.rds")

# Read in if closed R
CWM_overlap <- readRDS("CWM_overlap.rds")
```



### *Calculate the shift in CONTINUOUS traits*

```{r}
#rename columns
colnames(CWM_overlap) <- c("clutch_overlap", "bm_overlap", 
                           "GL_overlap", "foraging_overlap",
                           "diet_5cat_overlap", 
                           "Long", "Lat")

# bind the two dataframes - note there are some NAs because we removed absent latlongs in the CWM_overlap df
CWM_shift<- left_join(CWM_all, CWM_overlap)

#calculate the percentage difference
# CLUTCH SIZE
CWM_shift$clutch_diff <- ((CWM_shift$clutch_overlap - CWM_shift$clutch)/ CWM_shift$clutch) *100

# BODY MASS
CWM_shift$bm_diff <- ((CWM_shift$bm_overlap - CWM_shift$body_mass_median)/ CWM_shift$body_mass_median) *100

# GENERATION LENGTH
CWM_shift$GL_diff <- ((CWM_shift$GL_overlap - CWM_shift$GL)/ CWM_shift$GL) *100


```



### Assign marine region for each lat long

Marine regions from: 
Flanders Marine Institute (2021). Global Oceans and Seas, version 1. Available online at https://www.marineregions.org/. https://doi.org/10.14284/542.

```{r}

# Load packages
library("ggplot2");library("sf");library("rnaturalearth"); library("rnaturalearthdata")

# Read in marine regions
marineregions <- st_read("goas_v01.shp")

# Assign CRS
marineregions<- st_transform(marineregions, crs = 4326)

## Assign an ID column so that we can join polygons back to original data
CWM_shift$ID <- factor(1:38314)

# Extract CWM_shift coordinates
shift_coord <- CWM_shift %>% select(Lat, Long, ID)

# convert to spatial
shift_coord <- shift_coord %>% 
  st_as_sf(coords = c("Long", "Lat"),
           crs = 4326)

# check that the two objects have the same CRS
st_crs(marineregions) == st_crs(shift_coord)

# Error with the spherical geometry when joining below, so turn off the s2 processing first
sf::sf_use_s2(FALSE)

# locate for each point the nearest polygon
nearest_polygon <- st_join(shift_coord, marineregions,
                           join = st_nearest_feature)


# Select only the needed data
nearest_polygon <- nearest_polygon %>% rename(ocean = name) %>% select(ID, ocean)
nearest_polygon$geometry <- NULL

# Match the marine region to the CWM_shift dataframe
CWM_shift <- left_join(CWM_shift, nearest_polygon)
CWM_shift$ID <- NULL


```





### *Plot the shift in CONTINUOUS traits with GAMM and qGAM*

```{r}
library(mgcv); library(tidymv); library(EnvStats);library(ggplot2);library(qgam); library(MASS)

### GENERATION LENGTH ###

# build model
m_GL <- qgam(GL_diff~s(Lat, k=12), qu = 0.2, data = CWM_shift)

# returns a tibble with the predictions from all the terms in a gam model
m_GL_p <- predict_gam(m_GL)

# Identify extreme values to determine where to crop the figure
quantile(CWM_shift$GL_diff, c(.001, .999), na.rm = T)

# Add a heading with facet wrap
CWM_shift$gl <- "Generation Length"

gl <- ggplot(CWM_shift)+
  geom_point(aes(Lat, GL_diff), colour = "#56B4E9", shape = 1, alpha = 1/10, size = 2)+
  #geom_line(aes(Lat, fit), data = m_GL_p, colour = "black", size = 1)+
  coord_flip()+
  geom_hline(yintercept = 0, linetype="dashed", size = 1)+
  theme_bw(base_size = 20)+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(legend.position = "none") +
  labs(x = "", y = "Deviation in CWM (%)")+
 # ggtitle("Generation Length")+
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))+
  scale_y_continuous(limits = c(-21, 5.6))+
  scale_x_continuous(breaks = seq(-50, 50, 50),
                     labels = c(expression(50*degree*S), expression(0*degree*""), expression(50*degree*N)))+
   facet_wrap(~gl)+
  theme(strip.text.x = element_text(size = 22, color = "white", face = "bold"), 
          strip.background = element_rect(fill="black"))

  
ggsave("GL_shift.png")




### CLUTCH ###    

# build model
m_clutch <- gamm(clutch_diff~s(Lat), data = CWM_shift, method = "REML")

# returns a tibble with the predictions from all the terms in a gam model
m_clutch_p <- predict_gam(m_clutch$gam)

# Identify extreme values to determine where to crop the figure
quantile(CWM_shift$clutch_diff, c(.001, .999), na.rm = T)

# Add a heading with facet wrap
CWM_shift$cl <- "Clutch Size"

cl <- ggplot(CWM_shift)+
  geom_point(aes(Lat, clutch_diff), colour =  "#E69F00", shape = 1, alpha = 1/10, size = 2)+
  geom_line(aes(Lat, fit), data = m_clutch_p, colour = "black", size = 1)+
  coord_flip()+
  geom_hline(yintercept = 0, linetype="dashed", size = 1)+
  theme_bw(base_size = 20)+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(legend.position = "none") +
  labs(x = "Latitude", y = "Deviation in CWM (%)")+
  #ggtitle("Clutch Size")+
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))+
  scale_y_continuous(limits = c(-25, 55))+
  scale_x_continuous(breaks = seq(-50, 50, 50),
                     labels = c(expression(50*degree*S), expression(0*degree*""), expression(50*degree*N)))+
   facet_wrap(~cl)+
  theme(strip.text.x = element_text(size = 22, color = "white", face = "bold"), 
          strip.background = element_rect(fill="black"))
  
ggsave("clutch_shift.png")
  
  
  
### BODY MASS ###    


# build model
m_bm <- qgam(bm_diff~s(Lat, k=12), qu = 0.2, data = CWM_shift)

# returns a tibble with the predictions from all the terms in a gam model
m_bm_p <- predict_gam(m_bm)

# Identify extreme values to determine where to crop the figure
quantile(CWM_shift$bm_diff, c(.001, .999), na.rm = T)

# Add a heading with facet wrap
CWM_shift$bm <- "Body Mass"

bm <- ggplot(CWM_shift)+
  geom_point(aes(Lat, bm_diff), colour =  "#56B4E9", shape = 1, alpha = 1/10, size = 2)+
  geom_line(aes(Lat, fit), data = m_bm_p, colour = "black", size = 1)+
  coord_flip()+
  geom_hline(yintercept = 0, linetype="dashed", size = 1)+
  theme_bw(base_size = 20)+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(legend.position = "none") +
  labs(x = "", y = "Deviation in CWM (%)")+
  #ggtitle("Body Mass")+
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))+
 scale_y_continuous(limits = c(-39, 12))+
 scale_x_continuous(breaks = seq(-50, 50, 50),
                     labels = c(expression(50*degree*S), expression(0*degree*""), expression(50*degree*N)))+
  facet_wrap(~bm)+
  theme(strip.text.x = element_text(size = 22, color = "white", face = "bold"), 
          strip.background = element_rect(fill="black"))
  

ggsave("bodymass_shift.png")





# MAKE A PANEL PLOT 
ggarrange(cl,bm, gl, nrow = 1, ncol = 3, labels = c("A", "B", "C"), 
          font.label = list(size = 20, color = "black", face ="bold"))

ggsave("continuous.png", dpi = 300, width = 21, height = 7)

```



### *Calculate the shift in foraging guild (CATEGORICAL)*

```{r message=FALSE, error=FALSE, warning=FALSE}

library(tidyverse)

traits$binomial <- colnames(PAM)
PAM <- as.data.frame(PAM)
PAM_overlap <- as.data.frame(PAM_overlap)

#PAM2<- PAM_sens
#__________________________#
#     Surface              #  
#__________________________#


# 1. subset the species that are surface foragers from the trait dataset
surface <- traits[traits$foraging_guild %in% c("Surface"), ]

# 2. subset the surface foragers within the PAM
PAM_surface <- dplyr::select(PAM, surface$binomial)

# 3. calculate the frequency of of species across latitude from the PAM rows
PAM_surface$freq <- rowSums(PAM_surface) 

# 4. subset the surface foragers within the PAM_overlap 
# (locations where threatened species overlap with fishing activities have been removed)
PAM_surface_NT <- dplyr::select(PAM_overlap, surface$binomial)

# 5. calculate the frequency of of species across latitude from the PAM_overlap rows
PAM_surface_NT$freq <- rowSums(PAM_surface_NT)


#__________________________
# Diver
#__________________________

# 1. subset non_pelagic specialists
diver <- traits[traits$foraging_guild %in% c("Diver"), ]

# 2. subset PAM
PAM_diver <- dplyr::select(PAM, diver$binomial)

# 3. calculate the frequency of each PAM row
PAM_diver$freq <- rowSums(PAM_diver) 

# 4. subset PAM
PAM_diver_NT <- dplyr::select(PAM_overlap, diver$binomial)

# 5. calculate the frequency of each PAM row
PAM_diver_NT$freq <- rowSums(PAM_diver_NT)


#__________________________
# Ground
#__________________________

# 1. subset non_pelagic specialists
ground <- traits[traits$foraging_guild %in% c("Ground"), ]

# 2. subset PAM
PAM_ground <- dplyr::select(PAM, ground$binomial)

# 3. calculate the frequency of each PAM row
PAM_ground$freq <- rowSums(PAM_ground) 

# 4. subset PAM
PAM_ground_NT <- dplyr::select(PAM_overlap, ground$binomial)

# 5. calculate the frequency of each PAM row
PAM_ground_NT$freq <- rowSums(PAM_ground_NT)





#__________________________
# Generalist
#__________________________

# 1. subset non_pelagic specialists
general <- traits[traits$foraging_guild %in% c("Generalist"), ]

# 2. subset PAM
PAM_general <- dplyr::select(PAM, general$binomial)

# 3. calculate the frequency of each PAM row
PAM_general$freq <- rowSums(PAM_general) 

# 4. subset PAM
PAM_general_NT <- dplyr::select(PAM_overlap, general$binomial)

# 5. calculate the frequency of each PAM row
PAM_general_NT$freq <- rowSums(PAM_general_NT)



##_________
# Bind the frequencies into one dataframe in order to calculate the differences
# First, need to add the lat and longs back so that the two data types can be joined correctly
forage_full <- cbind(dplyr::select(CWM_all, Lat, Long),
                     PAM_surface$freq, PAM_diver$freq,PAM_ground$freq,PAM_general$freq)

forage_overlap <- cbind(dplyr::select(CWM_overlap, Lat, Long),
                        PAM_surface_NT$freq, PAM_diver_NT$freq,PAM_ground_NT$freq,PAM_general_NT$freq)

forage_diff <- left_join(forage_full, forage_overlap, by = c("Lat", "Long"))

colnames(forage_diff) <- c("Lat", "Long",
                           "surface", "diver", "ground","general",
                           "surface_NT", "diver_NT", "ground_NT","general_NT")


# calculate proportion of surface, diver vs. ground

forage_diff$surface_prop <- (forage_diff$surface/(forage_diff$surface + forage_diff$diver + forage_diff$ground + forage_diff$general))*100

forage_diff$diver_prop <- (forage_diff$diver/(forage_diff$surface + forage_diff$diver + forage_diff$ground + forage_diff$general))*100

forage_diff$ground_prop <- (forage_diff$ground/(forage_diff$surface + forage_diff$diver + forage_diff$ground + forage_diff$general))*100

forage_diff$general_prop <- (forage_diff$general/(forage_diff$surface + forage_diff$diver + forage_diff$ground + forage_diff$general))*100



# calculate proportion of  surface_NT, diver_NT vs. ground_NT
forage_diff$surface_NT_prop <- (forage_diff$surface_NT/(forage_diff$surface_NT + forage_diff$diver_NT + forage_diff$ground_NT + forage_diff$general_NT))*100

forage_diff$diver_NT_prop <- (forage_diff$diver_NT/(forage_diff$surface_NT + forage_diff$diver_NT + forage_diff$ground_NT + forage_diff$general_NT))*100

forage_diff$ground_NT_prop <- (forage_diff$ground_NT/(forage_diff$surface_NT + forage_diff$diver_NT + forage_diff$ground_NT + forage_diff$general_NT))*100

forage_diff$general_NT_prop <- (forage_diff$general_NT/(forage_diff$surface_NT + forage_diff$diver_NT + forage_diff$ground_NT + forage_diff$general_NT))*100





# calculate the difference for surface vs. surface_NT 
forage_diff$surface_diff <- forage_diff$surface_NT_prop - forage_diff$surface_prop

# calculate the difference for diver vs. diver_NT
forage_diff$diver_diff <- forage_diff$diver_NT_prop - forage_diff$diver_prop

# calculate the difference for ground vs. ground_NT
forage_diff$ground_diff <- forage_diff$ground_NT_prop - forage_diff$ground_prop

# calculate the difference for general vs. general_NT
forage_diff$general_diff <- forage_diff$general_NT_prop - forage_diff$general_prop


```


### *Plot the shift in foraging guild (CATEGORICAL) with GAMM*
```{r}

### SURFACE FORAGER ###

# build model
m_surface <- gamm(surface_diff~s(Lat), data = forage_diff)

# returns a tibble with the predictions from all the terms in a gam model
m_surface_p <- predict_gam(m_surface$gam)

# Identify extreme values to determine where to crop the figure
quantile(forage_diff$surface_diff, c(.01, .99), na.rm = T)


forage_diff$surf <- "Surface Forager"

p1<-ggplot(forage_diff)+
  geom_point(aes(Lat, surface_diff), colour = "#E69F00", shape = 1, alpha = 1/10, size = 2)+
  geom_line(aes(Lat, fit), data = m_surface_p, colour = "black", size = 1)+
  coord_flip()+
  geom_hline(yintercept = 0, linetype="dashed", size = 1)+
  theme_bw(base_size = 20)+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(legend.position = "none") +
  labs(x = "Latitude", y = "Deviation in CWM (%)")+
  #ggtitle("Surface Forager")+
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))+
 scale_y_continuous(limits = c(-50, 60))+
  scale_x_continuous(breaks = seq(-50, 50, 50),
                     labels = c(expression(50*degree*S), expression(0*degree*""), expression(50*degree*N)))+
  facet_wrap(~surf)+
  theme(strip.text.x = element_text(size = 22, color = "white", face = "bold"), 
          strip.background = element_rect(fill="black"))
  
ggsave("Surface_shift.png")



### DIVER ###

# build model
m_diver <- gamm(diver_diff~s(Lat), data = forage_diff)

# returns a tibble with the predictions from all the terms in a gam model
m_diver_p <- predict_gam(m_diver$gam)

# Identify extreme values to determine where to crop the figure
quantile(forage_diff$diver_diff, c(.01, .99), na.rm = T)

forage_diff$div <- "Diving Forager"

p2<- ggplot(forage_diff)+
  geom_point(aes(Lat, diver_diff), colour = "#56B4E9", shape = 1, alpha = 1/10, size = 2)+
  geom_line(aes(Lat, fit), data = m_diver_p, colour = "black", size = 1)+
  coord_flip()+
  geom_hline(yintercept = 0, linetype="dashed", size = 1)+
  theme_bw(base_size = 20)+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(legend.position = "none") +
  labs(x = "", y = "Deviation in CWM (%)")+
  #ggtitle("Diving Forager")+
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))+
  scale_y_continuous(limits = c(-57, 36))+
  scale_x_continuous(breaks = seq(-50, 50, 50),
                     labels = c(expression(50*degree*S), expression(0*degree*""), expression(50*degree*N)))+
    facet_wrap(~div)+
  theme(strip.text.x = element_text(size = 22, color = "white", face = "bold"), 
          strip.background = element_rect(fill="black"))
  
ggsave("Diver_shift.png")



### GENERALIST ###

# build model
m_general <- gamm(general_diff~s(Lat), data = forage_diff)

# returns a tibble with the predictions from all the terms in a gam model
m_general_p <- predict_gam(m_general$gam)


# Identify extreme values to determine where to crop the figure
quantile(forage_diff$general_diff, c(.01, .99), na.rm = T)

forage_diff$gen <- "Generalist Forager"

p3<- ggplot(forage_diff)+
  geom_point(aes(Lat, general_diff), colour = "#56B4E9", shape = 1, alpha = 1/10, size = 2)+
  geom_line(aes(Lat, fit), data = m_general_p, colour = "black", size = 1)+
  coord_flip()+
  geom_hline(yintercept = 0, linetype="dashed", size = 1)+
  theme_bw(base_size = 20)+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(legend.position = "none") +
  labs(x = "", y = "Deviation in CWM (%)")+
  #ggtitle("Generalist Forager")+
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))+
  scale_y_continuous(limits = c(-30, 37))+
  scale_x_continuous(breaks = seq(-50, 50, 50),
                     labels = c(expression(50*degree*S), expression(0*degree*""), expression(50*degree*N)))+
  facet_wrap(~gen)+
  theme(strip.text.x = element_text(size = 22, color = "white", face = "bold"), 
          strip.background = element_rect(fill="black"))
  
ggsave("Generalist_shift.png")


### GROUND ###

# build model
m_ground <- gamm(ground_diff~s(Lat), data = forage_diff)

# returns a tibble with the predictions from all the terms in a gam model
m_ground_p <- predict_gam(m_ground$gam)


# Identify extreme values to determine where to crop the figure
quantile(forage_diff$ground_diff, c(.01, .99), na.rm = T) # before = -35 & 32

forage_diff$gr <- "Ground Forager"

p4<- ggplot(forage_diff)+
  geom_point(aes(Lat, ground_diff), colour = "#E69F00", shape = 1, alpha = 1/10, size = 2)+
  geom_line(aes(Lat, fit), data = m_ground_p, colour = "black", size = 1)+
  coord_flip()+
  geom_hline(yintercept = 0, linetype="dashed", size = 1)+
  theme_bw(base_size = 20)+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(legend.position = "none") +
  labs(x = "Latitude", y = "Deviation in CWM (%)")+
  #ggtitle("Ground Forager")+
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))+
  scale_y_continuous(limits = c(-13, 75))+
  scale_x_continuous(breaks = seq(-50, 50, 50),
                     labels = c(expression(50*degree*S), expression(0*degree*""), expression(50*degree*N)))+
  facet_wrap(~gr)+
  theme(strip.text.x = element_text(size = 22, color = "white", face = "bold"), 
          strip.background = element_rect(fill="black"))
  
ggsave("Ground_shift.png")



ggarrange(p1, p2, p3, p4, 
          nrow = 2, ncol = 2, 
          labels = c("A", "B", "C", "D"), 
          font.label = list(size = 20, color = "black", face ="bold"))

ggsave("mix2.png", dpi = 300, width = 14, height = 14)

```


### *Calculate the shift in diet guild (CATEGORICAL)*

```{r message=FALSE, error=FALSE, warning=FALSE}

library(tidyverse)

traits$binomial <- colnames(PAM)
PAM <- as.data.frame(PAM)
PAM_overlap <- as.data.frame(PAM_overlap)

#PAM2<- PAM_sens
#__________________________#
#     Fish & carrion       #  
#__________________________#


# 1. subset the species that are vertebrate consumers from the trait dataset
vert <- traits[traits$diet_5cat %in% c("Fish and Carrion"), ]

# 2. subset the vertebrate consumers within the PAM
PAM_vert <- dplyr::select(PAM, vert$binomial)

# 3. calculate the frequency of of species across latitude from the PAM rows
PAM_vert$freq <- rowSums(PAM_vert) 

# 4. subset the vertebrate consumers within the PAM_overlap 
# (locations where threatened species overlap with fishing activities have been removed)
PAM_vert_NT <- dplyr::select(PAM_overlap, vert$binomial)

# 5. calculate the frequency of of species across latitude from the PAM_overlap rows
PAM_vert_NT$freq <- rowSums(PAM_vert_NT)


#__________________________
# Invertebrates
#__________________________

# 1. subset vertebrate consumers
invert <- traits[traits$diet_5cat %in% c("Invertebrates"), ]

# 2. subset PAM
PAM_invert <- dplyr::select(PAM, invert$binomial)

# 3. calculate the frequency of each PAM row
PAM_invert$freq <- rowSums(PAM_invert) 

# 4. subset PAM
PAM_invert_NT <- dplyr::select(PAM_overlap, invert$binomial)

# 5. calculate the frequency of each PAM row
PAM_invert_NT$freq <- rowSums(PAM_invert_NT)


#__________________________
# Omnivores
#__________________________

# 1. subset Omnivores
Omnivore <- traits[traits$diet_5cat %in% c("Omnivore"), ]

# 2. subset PAM
PAM_omniv <- dplyr::select(PAM, Omnivore$binomial)

# 3. calculate the frequency of each PAM row
PAM_omniv$freq <- rowSums(PAM_omniv) 

# 4. subset PAM
PAM_omniv_NT <- dplyr::select(PAM_overlap, Omnivore$binomial)

# 5. calculate the frequency of each PAM row
PAM_omniv_NT$freq <- rowSums(PAM_omniv_NT)



##_________
# Bind the frequencies into one dataframe in order to calculate the differences
# First, need to add the lat and longs back so that the two data types can be joined correctly
diet_full <- cbind(dplyr::select(CWM_all, Lat, Long),
                     PAM_vert$freq, PAM_invert$freq,PAM_omniv$freq)

diet_overlap <- cbind(dplyr::select(CWM_overlap, Lat, Long),
                        PAM_vert_NT$freq, PAM_invert_NT$freq,PAM_omniv_NT$freq)

diet_diff <- left_join(diet_full, diet_overlap, by = c("Lat", "Long"))

colnames(diet_diff) <- c("Lat", "Long",
                           "vert", "invert", "omnivore",
                           "vert_NT", "invert_NT", "omnivore_NT")



# calculate proportion

diet_diff$vert_prop <- (diet_diff$vert/
                               (diet_diff$vert + 
                                  diet_diff$invert + 
                                  diet_diff$omnivore))*100

diet_diff$invert_prop <- (diet_diff$invert/(diet_diff$vert + 
                                              diet_diff$invert + 
                                              diet_diff$omnivore))*100

diet_diff$omnivore_prop <- (diet_diff$omnivore/(diet_diff$vert + 
                                                  diet_diff$invert + 
                                                  diet_diff$omnivore))*100


# calculate proportion of  vert_NT, invert_NT vs. omnivore_NT
diet_diff$vert_NT_prop <- (diet_diff$vert_NT/(diet_diff$vert_NT + 
                                                diet_diff$invert_NT + 
                                                diet_diff$omnivore_NT))*100

diet_diff$invert_NT_prop <- (diet_diff$invert_NT/(diet_diff$vert_NT + 
                                                    diet_diff$invert_NT +
                                                    diet_diff$omnivore_NT))*100

diet_diff$omnivore_NT_prop <- (diet_diff$omnivore_NT/(diet_diff$vert_NT + 
                                                        diet_diff$invert_NT + 
                                                        diet_diff$omnivore_NT))*100





# calculate the difference for vert vs. vert_NT 
diet_diff$vert_diff <- diet_diff$vert_NT_prop - diet_diff$vert_prop

# calculate the difference for invert vs. invert_NT
diet_diff$invert_diff <- diet_diff$invert_NT_prop - diet_diff$invert_prop

# calculate the difference for omnivore vs. omnivore_NT
diet_diff$omnivore_diff <- diet_diff$omnivore_NT_prop - diet_diff$omnivore_prop


```









### *Plot the shift in diet guild (CATEGORICAL) with GAMM*
```{r}

### VERTEBRATE FORAGER ###

# build model
m_vert <- gamm(vert_diff~s(Lat), data = diet_diff)

# returns a tibble with the predictions from all the terms in a gam model
m_vert_p <- predict_gam(m_vert$gam)

# Identify extreme values to determine where to crop the figure
quantile(diet_diff$vert_diff, c(.01, .99), na.rm = T)


diet_diff$vertebrate <- "Vertebrate & Scavenger Diet"

vertebrates<-ggplot(diet_diff)+
  geom_point(aes(Lat, vert_diff), colour = "#E69F00", shape = 1, alpha = 1/10, size = 2)+
  geom_line(aes(Lat, fit), data = m_vert_p, colour = "black", size = 1)+
  coord_flip()+
  geom_hline(yintercept = 0, linetype="dashed", size = 1)+
  theme_bw(base_size = 20)+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(legend.position = "none") +
  labs(x = "Latitude", y = "Deviation in CWM (%)")+
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))+
 #scale_y_continuous(limits = c(-50, 60))+
  scale_x_continuous(breaks = seq(-50, 50, 50),
                     labels = c(expression(50*degree*S), expression(0*degree*""), expression(50*degree*N)))+
  facet_wrap(~vertebrate)+
  theme(strip.text.x = element_text(size = 22, color = "white", face = "bold"), 
          strip.background = element_rect(fill="black"))
  
ggsave("vertebrate_shift.png")



### INVERTEBRATES ###

# build model
m_invert <- gamm(invert_diff~s(Lat), data = diet_diff)

# returns a tibble with the predictions from all the terms in a gam model
m_invert_p <- predict_gam(m_invert$gam)

# Identify extreme values to determine where to crop the figure
quantile(diet_diff$invert_diff, c(.01, .99), na.rm = T)

diet_diff$invertebrate <- "Invertebrate Diet"

inverts<- ggplot(diet_diff)+
  geom_point(aes(Lat, invert_diff), colour = "#56B4E9", shape = 1, alpha = 1/10, size = 2)+
  geom_line(aes(Lat, fit), data = m_invert_p, colour = "black", size = 1)+
  coord_flip()+
  geom_hline(yintercept = 0, linetype="dashed", size = 1)+
  theme_bw(base_size = 20)+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(legend.position = "none") +
  labs(x = "", y = "Deviation in CWM (%)")+
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))+
  #scale_y_continuous(limits = c(-57, 36))+
  scale_x_continuous(breaks = seq(-50, 50, 50),
                     labels = c(expression(50*degree*S), expression(0*degree*""), expression(50*degree*N)))+
    facet_wrap(~invertebrate)+
  theme(strip.text.x = element_text(size = 22, color = "white", face = "bold"), 
          strip.background = element_rect(fill="black"))
  
ggsave("invertebrate_shift.png")



### Omnivore ###

# build model
m_omniv <- gamm(omnivore_diff~s(Lat), data = diet_diff)

# returns a tibble with the predictions from all the terms in a gam model
m_omniv_p <- predict_gam(m_omniv$gam)


# Identify extreme values to determine where to crop the figure
quantile(diet_diff$omnivore_diff, c(.01, .99), na.rm = T)

diet_diff$omnivores <- "Omnivore Diet"

Omnivores<- ggplot(diet_diff)+
  geom_point(aes(Lat, omnivore_diff), colour = "#56B4E9", shape = 1, alpha = 1/10, size = 2)+
  geom_line(aes(Lat, fit), data = m_omniv_p, colour = "black", size = 1)+
  coord_flip()+
  geom_hline(yintercept = 0, linetype="dashed", size = 1)+
  theme_bw(base_size = 20)+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(legend.position = "none") +
  labs(x = "", y = "Deviation in CWM (%)")+
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))+
  scale_y_continuous(limits = c(-30, 37))+
  scale_x_continuous(breaks = seq(-50, 50, 50),
                     labels = c(expression(50*degree*S), expression(0*degree*""), expression(50*degree*N)))+
  facet_wrap(~omnivores)+
  theme(strip.text.x = element_text(size = 22, color = "white", face = "bold"), 
          strip.background = element_rect(fill="black"))
  
ggsave("omnivore_shift.png")




ggarrange(vertebrates, inverts, Omnivores,
          nrow = 1, ncol = 3, 
          labels = c("A", "B", "C"), 
          font.label = list(size = 20, color = "black", face ="bold"))

ggsave("diet.png", dpi = 300, width = 21, height = 7)



ggarrange(cl, bm, gl,
          vertebrates, inverts, Omnivores,
          p1,p2,p3,p4,
          nrow = 4, ncol = 3, 
          labels = c("A", "B", "C",
                     "D", "E", "F",
                     "G","H","I","J"), 
          font.label = list(size = 20, color = "black", face ="bold"))


ggsave("allShifts.png", dpi = 300, width = 21, height = 25)
```




